---
title: "NFL BIG DATA BOWL"
output: html_document
---

#necessary packages
```{r}
library(tidyverse)
library(lubridate) 
library(beeswarm) 
library(gganimate)
library(ggridges)  
library(tidyr)
library(cowplot)
```

#to animate play
```{r}
play <- tracking2020 %>% filter(FinalID == 2021010301116)


## General field boundaries
xmin <- 0
xmax <- 160/3
hash.right <- 38.35
hash.left <- 12
hash.width <- 3.3


## Specific boundaries for a given play
ymin <- max(round(min(play$x, na.rm = TRUE) - 10, -1), 0)
ymax <- min(round(max(play$x, na.rm = TRUE) + 10, -1), 120)
df_hash <- expand.grid(x = c(0, 23.36667, 29.96667, xmax), y = (10:110))
df_hash <- df_hash %>% filter(!(floor(y %% 5) == 0))
df_hash <- df_hash %>% filter(y < ymax, y > ymin)

animate_play <- ggplot() +
  scale_size_manual(values = c(6, 4, 6), guide = FALSE) + 
  scale_shape_manual(values = c(21, 16, 21), guide = FALSE) +
  scale_fill_manual(values = c("#e31837", "#654321", "#002244"), guide = FALSE) + 
  scale_colour_manual(values = c("black", "#654321", "#c60c30"), guide = FALSE) + 
  annotate("text", x = df_hash$x[df_hash$x < 55/2], 
           y = df_hash$y[df_hash$x < 55/2], label = "_", hjust = 0, vjust = -0.2) + 
  annotate("text", x = df_hash$x[df_hash$x > 55/2], 
           y = df_hash$y[df_hash$x > 55/2], label = "_", hjust = 1, vjust = -0.2) + 
  annotate("segment", x = xmin, 
           y = seq(max(10, ymin), min(ymax, 110), by = 5), 
           xend =  xmax, 
           yend = seq(max(10, ymin), min(ymax, 110), by = 5)) + 
  annotate("text", x = rep(hash.left, 11), y = seq(10, 110, by = 10), 
           label = c("G   ", seq(10, 50, by = 10), rev(seq(10, 40, by = 10)), "   G"), 
           angle = 270, size = 4) + 
  annotate("text", x = rep((xmax - hash.left), 11), y = seq(10, 110, by = 10), 
           label = c("   G", seq(10, 50, by = 10), rev(seq(10, 40, by = 10)), "G   "), 
           angle = 90, size = 4) + 
  annotate("segment", x = c(xmin, xmin, xmax, xmax), 
           y = c(ymin, ymax, ymax, ymin), 
           xend = c(xmin, xmax, xmax, xmin), 
           yend = c(ymax, ymax, ymin, ymin), colour = "black") + 
  geom_point(data = play, aes(x = (xmax-y), y = x, shape = team,
                              fill = team, group = nflId, size = team, colour = team), alpha = 0.7) + 
  geom_text(data = play, aes(x = (xmax-y), y = x, label = jerseyNumber), colour = "white", 
            vjust = 0.36, size = 3.5) + 
  ylim(ymin, ymax) + 
  coord_fixed() +  
  theme_nothing() + 
  transition_time(frameId)  +
  ease_aes('linear') + 
  NULL

## Ensure timing of play matches 10 frames-per-second
play.length.ex <- length(unique(play$frameId))

#to animate play
animate(animate_play, fps = 10, nframe = play.length.ex)
```

#loading in tracking data
```{r}
tracking2020 <- read_csv('tracking2020.csv')
tracking2019 <- read_csv('tracking2019.csv')
tracking2018 <- read_csv('tracking2018.csv')

tracking2020$FinalID <- paste0(tracking2020$gameId, tracking2020$playId)
tracking2019$FinalID <- paste0(tracking2019$gameId, tracking2019$playId)
tracking2018$FinalID <- paste0(tracking2018$gameId, tracking2018$playId)

```

#loading in other data
```{r}
games <- read_csv('games.csv')
PFF <- read_csv('PFFScoutingData.csv')
players <- read_csv('players.csv')
plays <- read_csv('plays.csv')
```

#keeping punt plays
```{r}
plays_punt <- plays %>% filter(specialTeamsPlayType == 'Punt')

plays_punt$FinalID <- paste0(plays_punt$gameId, plays_punt$playId)
```


#2020 Gunner and Vises Max Speeds
```{r}
#using df_tracking to merge to jersey numbers
df_jerseyMap <- tracking2020 %>% 
  
  #selecting variables of interest
  select(gameId, team, jerseyNumber, nflId) %>%
  
  #dropping duplicates - jersey # is constant throughout game
  distinct() %>%
  
  #joining to games
  inner_join(games, by = 'gameId') %>%
  
  #getting name of team
  mutate(team = ifelse(team == 'home', homeTeamAbbr, visitorTeamAbbr),
         
         #adjusting jersey number so that it includes 0 when < 10
         jerseyNumber = ifelse(jerseyNumber < 10,
                               paste0('0', jerseyNumber),
                               as.character(jerseyNumber)),
         
         #getting team and jersey
         teamJersey = paste0(team, ' ', jerseyNumber)) %>%
  
  #map to merge nflId to teamJersey
  select(gameId, nflId, teamJersey) %>%
  
  arrange(gameId, nflId, teamJersey)
```

Gunners 2020
```{r}
###Gunner
df_PFF_specialGunners <- PFF %>%
  
  #splitting into a column for each special teams gunner
  separate(gunners, sep = '; ', into = c("teamJersey1",
                                                      "teamJersey2",
                                                      "teamJersey3",
                                                      "teamJersey4",
                                                      "teamJersey5",
                                                      "teamJersey6",
                                                      "teamJersey7",
                                                      "teamJersey8",
                                                      "teamJersey9",
                                                      "teamJersey10",
                                                      "teamJersey11")) %>%
  
  
  #selecting jersey numbers for each team
  select(gameId, playId, teamJersey1, teamJersey2,
         teamJersey3, teamJersey4, teamJersey5,
         teamJersey6, teamJersey7, teamJersey8,
         teamJersey9, teamJersey10, teamJersey11) %>%
  
  #gathering data
  gather(key = 'Type', value = 'teamJersey', -gameId, -playId) %>%
  
  #dropping NA rows
  drop_na() %>%
  
  #joining to jersey map
  inner_join(df_jerseyMap, by = c("gameId", 'teamJersey')) %>%
  
  #selecting variables of interest
  select(gameId, playId, nflId) %>%
  
  arrange(gameId, playId, nflId)
```

Vises 2020
```{r}
###Vises

df_PFF_specialVises <- PFF %>%
  
  #splitting into a column for each special teams vise
  separate(vises, sep = '; ', into = c("teamJersey1",
                                         "teamJersey2",
                                         "teamJersey3",
                                         "teamJersey4",
                                         "teamJersey5",
                                         "teamJersey6",
                                         "teamJersey7",
                                         "teamJersey8",
                                         "teamJersey9",
                                         "teamJersey10",
                                         "teamJersey11")) %>%
  
  
  #selecting jersey numbers for each team
  select(gameId, playId, teamJersey1, teamJersey2,
         teamJersey3, teamJersey4, teamJersey5,
         teamJersey6, teamJersey7, teamJersey8,
         teamJersey9, teamJersey10, teamJersey11) %>%
  
  #gathering data
  gather(key = 'Type', value = 'teamJersey', -gameId, -playId) %>%
  
  #dropping NA rows
  drop_na() %>%
  
  #joining to jersey map
  inner_join(df_jerseyMap, by = c("gameId", 'teamJersey')) %>%
  
  #selecting variables of interest
  select(gameId, playId, nflId) %>%
  
  arrange(gameId, playId, nflId)
```


finding out the average time of when ball lands on punts
```{r}
#events are punt_received, punt_land, and fair_catch

tracking20test <- tracking2020 %>% select(gameId, playId, nflId, event, frameId) %>% distinct() %>% filter(event == c('punt_received','punt_land' ,'fair_catch'))
tracking19test <- tracking2019 %>% select(gameId, playId, nflId, event, frameId) %>% distinct() %>% filter(event == c("punt_land", "punt_received","fair_catch"))
tracking18test <- tracking2018 %>% select(gameId, playId, nflId, event, frameId) %>% distinct() %>% filter(event == c('punt_received', 'punt_land','fair_catch'))

frame_avg <- rbind(tracking20test,tracking19test,tracking18test)

summary(frame_avg)

frame_cut_off <- 80 + (1.5*6)
frame_cut_off
```

getting max speeds for gunners before start of punt return 2020
```{r}
df_maxSpeeds <- tracking2020 %>%
  
  #joining games
  inner_join(games, by = 'gameId') %>%
  
  #only including gunners
  inner_join(df_PFF_specialGunners, by = c("gameId", 'playId', 'nflId'))  %>%
  
  #joining punt plays
  inner_join(plays_punt, by = c('gameId', 'playId')) %>%
  
  #removing the kicker from the tracking data
  filter(kickerId != nflId,
         
         #player is on home team and kicking team is home
         ((team == 'home') & (possessionTeam == homeTeamAbbr)) |
           ((team == 'away') & (possessionTeam == visitorTeamAbbr))  ) %>%
  
  #ungrouping
  ungroup() %>%
  
  #grouping by gameId, playId and nflId
  group_by(gameId, playId, nflId) %>%
  
  #filtering for first 89 observations
  filter(row_number() <= 89) %>%
  
  #calculating max speed for given play / player
  summarize(maxSpeed = max(s, na.rm = T), .groups = 'keep' ) %>%
  
  #ungrouping
  ungroup()

```


getting max speeds for vises before start of punt return 2020
```{r}
df_maxSpeeds_vises <- tracking2020 %>%
  
  #joining games
  inner_join(games, by = 'gameId') %>%
  
  #only including gunners
  inner_join(df_PFF_specialVises, by = c("gameId", 'playId', 'nflId'))  %>%
  
  #joining punt plays
  inner_join(plays_punt, by = c('gameId', 'playId')) %>%
  
  filter(kickerId != nflId) %>%
         #((team == "home") & (possessionTeam == homeTeamAbbr)) |
           #((team == "away") & (possessionTeam == visitorTeamAbbr))) 
  
  #ungrouping
  ungroup() %>%
  
  #grouping by gameId, playId and nflId
  group_by(gameId, playId, nflId) %>%
  
  #filtering for first 89 observations
  filter(row_number() <= 89) %>%
  
  #calculating max speed for given play / player
  summarize(maxSpeed = max(s, na.rm = T), .groups = 'keep') %>%
  
  #ungrouping
  ungroup()

```
#2019 Gunner and Vises Max Speeds
```{r}
#using df_tracking to merge to jersey numbers
df_jerseyMap19 <- tracking2019 %>% 
  
  #selecting variables of interest
  select(gameId, team, jerseyNumber, nflId) %>%
  
  #dropping duplicates - jersey # is constant throughout game
  distinct() %>%
  
  #joining to games
  inner_join(games, by = 'gameId') %>%
  
  #getting name of team
  mutate(team = ifelse(team == 'home', homeTeamAbbr, visitorTeamAbbr),
         
         #adjusting jersey number so that it includes 0 when < 10
         jerseyNumber = ifelse(jerseyNumber < 10,
                               paste0('0', jerseyNumber),
                               as.character(jerseyNumber)),
         
         #getting team and jersey
         teamJersey = paste0(team, ' ', jerseyNumber)) %>%
  
  #map to merge nflId to teamJersey
  select(gameId, nflId, teamJersey) %>%
  
  arrange(gameId, nflId, teamJersey)
```

Gunners 2019
```{r}
###Gunner
df_PFF_specialGunners19 <- PFF %>%
  
  #splitting into a column for each special teams gunner
  separate(gunners, sep = '; ', into = c("teamJersey1",
                                                      "teamJersey2",
                                                      "teamJersey3",
                                                      "teamJersey4",
                                                      "teamJersey5",
                                                      "teamJersey6",
                                                      "teamJersey7",
                                                      "teamJersey8",
                                                      "teamJersey9",
                                                      "teamJersey10",
                                                      "teamJersey11")) %>%
  
  
  #selecting jersey numbers for each team
  select(gameId, playId, teamJersey1, teamJersey2,
         teamJersey3, teamJersey4, teamJersey5,
         teamJersey6, teamJersey7, teamJersey8,
         teamJersey9, teamJersey10, teamJersey11) %>%
  
  #gathering data
  gather(key = 'Type', value = 'teamJersey', -gameId, -playId) %>%
  
  #dropping NA rows
  drop_na() %>%
  
  #joining to jersey map
  inner_join(df_jerseyMap19, by = c("gameId", 'teamJersey')) %>%
  
  #selecting variables of interest
  select(gameId, playId, nflId) %>%
  
  arrange(gameId, playId, nflId)
```

Vises 2019
```{r}
###Vises

df_PFF_specialVises19 <- PFF %>%
  
  #splitting into a column for each special teams vise
  separate(vises, sep = '; ', into = c("teamJersey1",
                                         "teamJersey2",
                                         "teamJersey3",
                                         "teamJersey4",
                                         "teamJersey5",
                                         "teamJersey6",
                                         "teamJersey7",
                                         "teamJersey8",
                                         "teamJersey9",
                                         "teamJersey10",
                                         "teamJersey11")) %>%
  
  
  #selecting jersey numbers for each team
  select(gameId, playId, teamJersey1, teamJersey2,
         teamJersey3, teamJersey4, teamJersey5,
         teamJersey6, teamJersey7, teamJersey8,
         teamJersey9, teamJersey10, teamJersey11) %>%
  
  #gathering data
  gather(key = 'Type', value = 'teamJersey', -gameId, -playId) %>%
  
  #dropping NA rows
  drop_na() %>%
  
  #joining to jersey map
  inner_join(df_jerseyMap19, by = c("gameId", 'teamJersey')) %>%
  
  #selecting variables of interest
  select(gameId, playId, nflId) %>%
  
  arrange(gameId, playId, nflId)
```

getting max speeds for gunners before start of punt return 2019
```{r}
df_maxSpeeds19 <- tracking2019 %>%
  
  #joining games
  inner_join(games, by = 'gameId') %>%
  
  #only including gunners
  inner_join(df_PFF_specialGunners19, by = c("gameId", 'playId', 'nflId'))  %>%
  
  #joining punt plays
  inner_join(plays_punt, by = c('gameId', 'playId')) %>%
  
  #removing the kicker from the tracking data
  filter(kickerId != nflId,
         
         #player is on home team and kicking team is home
         ((team == 'home') & (possessionTeam == homeTeamAbbr)) |
           ((team == 'away') & (possessionTeam == visitorTeamAbbr))  ) %>%

  #ungrouping
  ungroup() %>%
  
  #grouping by gameId, playId and nflId
  group_by(gameId, playId, nflId) %>%
  
  #filtering for first 89 observations
  filter(row_number() <= 89) %>%
  
  #calculating max speed for given play / player
  summarize(maxSpeed = max(s, na.rm = T), .groups = 'keep') %>%
  
  #ungrouping
  ungroup()
```

getting max speeds for vises before start of punt return 2019
```{r}
df_maxSpeeds_vises19 <- tracking2019 %>%
  
  #joining games
  inner_join(games, by = 'gameId') %>%
  
  #only including gunners
  inner_join(df_PFF_specialVises19, by = c("gameId", 'playId', 'nflId'))  %>%
  
  #joining punt plays
  inner_join(plays_punt, by = c('gameId', 'playId')) %>%
  
  filter(kickerId != nflId) %>%
         #((team == "home") & (possessionTeam == homeTeamAbbr)) |
           #((team == "away") & (possessionTeam == visitorTeamAbbr))) 
  
  #ungrouping
  ungroup() %>%
  
  #grouping by gameId, playId and nflId
  group_by(gameId, playId, nflId) %>%
  
  #filtering for first 89 observations
  filter(row_number() <= 89) %>%
  
  #calculating max speed for given play / player
  summarize(maxSpeed = max(s, na.rm = T), .groups = 'keep') %>%
  
  #ungrouping
  ungroup()

```


#2018 Gunner and Vises Max Speeds
```{r}
#using df_tracking to merge to jersey numbers
df_jerseyMap18 <- tracking2018 %>% 
  
  #selecting variables of interest
  select(gameId, team, jerseyNumber, nflId) %>%
  
  #dropping duplicates - jersey # is constant throughout game
  distinct() %>%
  
  #joining to games
  inner_join(games, by = 'gameId') %>%
  
  #getting name of team
  mutate(team = ifelse(team == 'home', homeTeamAbbr, visitorTeamAbbr),
         
         #adjusting jersey number so that it includes 0 when < 10
         jerseyNumber = ifelse(jerseyNumber < 10,
                               paste0('0', jerseyNumber),
                               as.character(jerseyNumber)),
         
         #getting team and jersey
         teamJersey = paste0(team, ' ', jerseyNumber)) %>%
  
  #map to merge nflId to teamJersey
  select(gameId, nflId, teamJersey) %>%
  
  arrange(gameId, nflId, teamJersey)
```

Gunners 2018
```{r}
###Gunner
df_PFF_specialGunners18 <- PFF %>%
  
  #splitting into a column for each special teams gunner
  separate(gunners, sep = '; ', into = c("teamJersey1",
                                                      "teamJersey2",
                                                      "teamJersey3",
                                                      "teamJersey4",
                                                      "teamJersey5",
                                                      "teamJersey6",
                                                      "teamJersey7",
                                                      "teamJersey8",
                                                      "teamJersey9",
                                                      "teamJersey10",
                                                      "teamJersey11")) %>%
  
  
  #selecting jersey numbers for each team
  select(gameId, playId, teamJersey1, teamJersey2,
         teamJersey3, teamJersey4, teamJersey5,
         teamJersey6, teamJersey7, teamJersey8,
         teamJersey9, teamJersey10, teamJersey11) %>%
  
  #gathering data
  gather(key = 'Type', value = 'teamJersey', -gameId, -playId) %>%
  
  #dropping NA rows
  drop_na() %>%
  
  #joining to jersey map
  inner_join(df_jerseyMap18, by = c("gameId", 'teamJersey')) %>%
  
  #selecting variables of interest
  select(gameId, playId, nflId) %>%
  
  arrange(gameId, playId, nflId)
```

Vises 2018
```{r}
###Vises

df_PFF_specialVises18 <- PFF %>%
  
  #splitting into a column for each special teams vise
  separate(vises, sep = '; ', into = c("teamJersey1",
                                         "teamJersey2",
                                         "teamJersey3",
                                         "teamJersey4",
                                         "teamJersey5",
                                         "teamJersey6",
                                         "teamJersey7",
                                         "teamJersey8",
                                         "teamJersey9",
                                         "teamJersey10",
                                         "teamJersey11")) %>%
  
  
  #selecting jersey numbers for each team
  select(gameId, playId, teamJersey1, teamJersey2,
         teamJersey3, teamJersey4, teamJersey5,
         teamJersey6, teamJersey7, teamJersey8,
         teamJersey9, teamJersey10, teamJersey11) %>%
  
  #gathering data
  gather(key = 'Type', value = 'teamJersey', -gameId, -playId) %>%
  
  #dropping NA rows
  drop_na() %>%
  
  #joining to jersey map
  inner_join(df_jerseyMap18, by = c("gameId", 'teamJersey')) %>%
  
  #selecting variables of interest
  select(gameId, playId, nflId) %>%
  
  arrange(gameId, playId, nflId)
```

getting max speeds for gunners before start of punt return 2018
```{r}
df_maxSpeeds18 <- tracking2018 %>%
  
  #joining games
  inner_join(games, by = 'gameId') %>%
  
  #only including gunners
  inner_join(df_PFF_specialGunners18, df_PFF_specialVises18, by = c("gameId", 'playId', 'nflId'))  %>%
  
  #joining punt plays
  inner_join(plays_punt, by = c('gameId', 'playId')) %>%
  
  #removing the kicker from the tracking data
  filter(kickerId != nflId,
         
         #player is on home team and kicking team is home
         ((team == 'home') & (possessionTeam == homeTeamAbbr)) |
           ((team == 'away') & (possessionTeam == visitorTeamAbbr))  ) %>%
  
  #ungrouping
  ungroup() %>%
  
  #grouping by gameId, playId and nflId
  group_by(gameId, playId, nflId) %>%
  
  #filtering for first 89 observations
  filter(row_number() <= 89) %>%
  
  #calculating max speed for given play / player
  summarize(maxSpeed = max(s, na.rm = T), .groups = 'keep') %>%
  
  #ungrouping
  ungroup()
```

getting max speeds for vises before start of punt return 2018
```{r}
df_maxSpeeds_vises18 <- tracking2018 %>%
  
  #joining games
  inner_join(games, by = 'gameId') %>%
  
  #only including gunners
  inner_join(df_PFF_specialVises18, by = c("gameId", 'playId', 'nflId'))  %>%
  
  #joining punt plays
  inner_join(plays_punt, by = c('gameId', 'playId')) %>%
  
  filter(kickerId != nflId) %>%
         #((team == "home") & (possessionTeam == homeTeamAbbr)) |
           #((team == "away") & (possessionTeam == visitorTeamAbbr))) 
  
  #ungrouping
  ungroup() %>%
  
  #grouping by gameId, playId and nflId
  group_by(gameId, playId, nflId) %>%
  
  #filtering for first 89 observations
  filter(row_number() <= 89) %>%
  
  #calculating max speed for given play / player
  summarize(maxSpeed = max(s, na.rm = T), .groups = 'keep') %>%
  
  #ungrouping
  ungroup()

```

#Combining datasets to create a dataframe with the max speed of every of every gunner on every punt return for the years 2018, 2019, and 2021

```{r}
gunner_speeds <- rbind(df_maxSpeeds, df_maxSpeeds19, df_maxSpeeds18)

vise_speeds <- rbind(df_maxSpeeds_vises, df_maxSpeeds_vises19, df_maxSpeeds_vises18)


#write_csv(gunner_speeds, 'gunner_speeds.csv')

#write_csv(vise_speeds, 'vise_speeds.csv')

```


#Comparing Gunner vs Vises Location

merging 2018 data frames
```{r}

gunner18 <- tracking2018 %>%
  
  #joining games
  inner_join(games, by = 'gameId') %>%
  
  #only including gunners and vises
  inner_join(df_PFF_specialGunners18, by = c("gameId", 'playId', 'nflId'))  %>%
  
  #joining punt plays
  inner_join(plays_punt, by = c('gameId', 'playId')) %>%
  
  #removing the kicker from the tracking data
  filter(kickerId != nflId) %>%
  
  #ungrouping
  ungroup() %>%
  
  #filter to just looking at locations at snap of the ball
  
  filter(event == 'ball_snap')

gunner18$gunner <- 1

vise18 <- tracking2018 %>%
  
  #joining games
  inner_join(games, by = 'gameId') %>%
  
  #only including gunners and vises
  inner_join(df_PFF_specialVises18, by = c("gameId", 'playId', 'nflId'))  %>%
  
  #joining punt plays
  inner_join(plays_punt, by = c('gameId', 'playId')) %>%
  
  #removing the kicker from the tracking data
  filter(kickerId != nflId) %>%
  
  #ungrouping
  ungroup() %>%
  
  #filter to just looking at locations at snap of the ball
  
  filter(event == 'ball_snap')

vise18$gunner <- 0

gunner_vise18 <- rbind(gunner18, vise18)

```


merging 2019 data frames
```{r}

gunner19 <- tracking2019 %>%
  
  #joining games
  inner_join(games, by = 'gameId') %>%
  
  #only including gunners and vises
  inner_join(df_PFF_specialGunners19, by = c("gameId", 'playId', 'nflId'))  %>%
  
  #joining punt plays
  inner_join(plays_punt, by = c('gameId', 'playId')) %>%
  
  #removing the kicker from the tracking data
  filter(kickerId != nflId) %>%
  
  #ungrouping
  ungroup() %>%
  
  #filter to just looking at locations at snap of the ball
  
  filter(event == 'ball_snap')

gunner19$gunner <- 1

vise19 <- tracking2019 %>%
  
  #joining games
  inner_join(games, by = 'gameId') %>%
  
  #only including gunners and vises
  inner_join(df_PFF_specialVises19, by = c("gameId", 'playId', 'nflId'))  %>%
  
  #joining punt plays
  inner_join(plays_punt, by = c('gameId', 'playId')) %>%
  
  #removing the kicker from the tracking data
  filter(kickerId != nflId) %>%
  
  #ungrouping
  ungroup() %>%
  
  #filter to just looking at locations at snap of the ball
  
  filter(event == 'ball_snap')

vise19$gunner <- 0


gunner_vise19 <- rbind(gunner19, vise19)

```


merging 2020 data frames
```{r}

gunner20 <- tracking2020 %>%
  
  #joining games
  inner_join(games, by = 'gameId') %>%
  
  #only including gunners and vises
  inner_join(df_PFF_specialGunners, by = c("gameId", 'playId', 'nflId'))  %>%
  
  #joining punt plays
  inner_join(plays_punt, by = c('gameId', 'playId')) %>%
  
  #removing the kicker from the tracking data
  filter(kickerId != nflId) %>%
  
  #ungrouping
  ungroup() %>%
  
  #filter to just looking at locations at snap of the ball
  
  filter(event == 'ball_snap')

gunner20$gunner <- 1

vise20 <- tracking2020 %>%
  
  #joining games
  inner_join(games, by = 'gameId') %>%
  
  #only including gunners and vises
  inner_join(df_PFF_specialVises, by = c("gameId", 'playId', 'nflId'))  %>%
  
  #joining punt plays
  inner_join(plays_punt, by = c('gameId', 'playId')) %>%
  
  #removing the kicker from the tracking data
  filter(kickerId != nflId) %>%
  
  #ungrouping
  ungroup() %>%
  
  #filter to just looking at locations at snap of the ball
  
  filter(event == 'ball_snap')

vise20$gunner <- 0

gunner_vise20 <- rbind(gunner20, vise20)

```

gunner and vise coordinates at snap of ball
```{r}

gunner_vise_coord <- rbind(gunner_vise20, gunner_vise19, gunner_vise18)
#write_csv(gunner_vise_coord, "gunner_vise_coord.csv")
```


#showing if there is a double team on a gunner at snap of ball

```{r}
gunner_vise_coord <-
  gunner_vise_coord %>% 
  group_by(FinalID.x) %>% 
  mutate(double_team = if_else(sum(gunner == 1) < sum(gunner == 0),1,0)) %>%
  ungroup()

```

#looking at play by play data for gunners
```{r}
pff_pbp_gunner <- gunner_speeds %>% left_join(plays_punt, c('gameId', 'playId'))
```

#merging doule team data

