---
title: "EDA on what Chararteristics Correlate with Income"
output:
  pdf_document: 
    toc: true
    toc_depth: 2
  html_document:
    df_print: paged
---

# Introduction
Team Members:

Alex Pesantez\

In this document, we survey the data from NLSY '79 and perform exploratory data analysis on the effects on income of combinations of other variables provided.

Our analysis will be to study the effects of education, race, gender and height on levels of income from 1981 to 2014. To begin, we will explore the effects of gender and height as it pertains to inocome.


## Study of Income
When looking at the unique years, you can see that the years vary from 1982 to 2014, which means that we have income data for that time frame. When looking at the histogram of the income data, you can clearly see that the data is skewed to the right with some major outliers. While there are definitely some outliers, in some of the analyses these won't be removed because not everything is being modeled and we just want to see the general sense of the trends between the variables, and in real life there are people that have drastically more or less income than others which we want to demonstrate in the analysis.

```{r}
options(scipen = 100)
library(tidyverse)
load("income_data_nlsy79.RData")
glimpse(income_data_nlsy79)
unique(income_data_nlsy79$year)
ggplot(income_data_nlsy79, aes(income)) + geom_histogram() + ggtitle("Histogram of Income 1982-2014") +
  theme(plot.title = element_text(hjust = 0.5,face = "bold"))
  
unique(filter(
  income_data_nlsy79,
  !is.na(income)
  )$year)
```
## Study of Gender and Height
### Physical Characteristics and Study of Gender and Height

```{r}
load("physical_data_nlsy79.RData")
unique(physical_data_nlsy79$year)
sort(unique(physical_data_nlsy79$height))
unique(physical_data_nlsy79$sex)
unique(filter(
  physical_data_nlsy79,
  !is.na(height)
  )$year)
ggplot(physical_data_nlsy79, aes(height)) + geom_histogram() + ggtitle("Histogram of Height 1981-2014") +
  theme(plot.title = element_text(hjust = 0.5,face = "bold"))
```

```{r}
unique(filter(
  physical_data_nlsy79,
  !is.na(sex)
  )$year)
ggplot(physical_data_nlsy79, aes(sex)) + geom_bar() + ggtitle("Histogram of Gender 1981-2014") +
  theme(plot.title = element_text(hjust = 0.5,face = "bold"))
```

```{r}
phys_income_all <- income_data_nlsy79 %>%
  inner_join(physical_data_nlsy79)
ggplot(phys_income_all, aes(sex)) + geom_bar() + ggtitle("Histogram of Gender 1982-2014") +
  theme(plot.title = element_text(hjust = 0.5,face = "bold"))
unique((physical_data_nlsy79$height %in% phys_income_all$height))
```

When looking at height and gender, we first wanted to see the range and different values there were for each variable. When looking at the unique values for each of the variables, you can see that there are only two genders observed, which are Male and Female, and that the height ranges from 48 inches to 91 inches tall. When looking at the years observed for gender, the values were from 1981 until 2014. When looking at the years for height, the data that was collected was only from 1981-1985 and 2006-2014. Since the years vary between all three variables, when we combine all three variables into a dataset, the only years that will be available are the years that contain the height data. When looking at the bar charts for the amount of each gender in the original data set, you can see that there were slightly more males than female. The bar chart for the combined data set shows the same story, with again more male than female. Then when observing the distribution for height you can see that there is a clear peak between 65 and 70 inches and that the range again varies from lower than 50 inches to greater than 80 inches.

### Exploring the effects of Gender on Income

```{r}
nrow(phys_income_all %>% filter(sex == "male" & !is.na(income)))
nrow(phys_income_all %>% filter(sex == "female" & !is.na(income)))
ggplot(
  data = phys_income_all,
  aes(x = sex, y = income)
) + geom_boxplot()
ggplot(data = phys_income_all, aes(income)) + geom_histogram() + facet_wrap(~sex)
```

When observing the effect of gender on income with a box plot, you can see that there are again some clear outliers. Within this analysis, they will not be removed due to the fact that again we are just observing the trends to get a general sense of how income and gender have related to one another in the past 20 to 30 years. In the box plot, you can see that the mean for male's is greater than the mean for female's which means in the past 30 years in general, males have had more income than females.\ 
The next thing that we observed was the distributions of income for each gender. When looking at both of the histograms, you can see that there is again right skewness, however there are higher counts of higher incomes for males when compared to females, which makes sense given that they have a higher mean income within the data set. Another clear difference between the two histograms is that female's had a lot more observations of zero income when compared to the males. This could be due to the fact that in earlier years, women were stereotyped as "stay-at-home moms", and it could've been harder for them to get a well payed job back then.

### Exploring the effects of Gender and Year on Income
The next thing we wanted to look at was the effect of gender and year on income. In order to do this, we made a graph which was broken up between female and male with year on the x-axis and income on the y-axis. When looking at this graph, you can see that for both genders as time went on the higher the income there was for each gender. This could just be due to inflation but still interesting to theorize. Another observation that can be made from this, is that the slope for the male's is greater than the slope for female's. This could mean that male's have a higher ceiling or potential when searching for a job with a high income than a female.

```{r}
ggplot(data = phys_income_all, aes(year,income)) +
  geom_jitter() + geom_smooth(method = "lm") + facet_wrap(~sex)
```

### Exploring the combined effects of Gender and Height on Income
```{r}
ggplot(data = phys_income_all, aes(height,income)) + geom_jitter() +
  geom_smooth(method = "lm") + facet_wrap(~sex)
```
The next thing explored was the combined effects of gender and height on income. Both of these graphs show a general trend that no matter what gender you are, the taller you the more income you will get. You can see this by the positive slopes for both male's and female's. Another takeaway you can see when looking at both graphs is that the slope for the male's is clearly greater than the slope for the female's, this means that being taller as a man is more impactful on income than it is being a taller female. This could be due to the fact people could find taller male's more intimidating than taller females which leads to them getting a higher income. This could also just be due to the fact that Male's have just made a lot more income in the data set skewing the slope to be greater just due to the fact of them having a greater income in general.

### Mean income by gender
Here you can see that male's on average made 12189.38 more than female's in the data observed.
```{r}
filter(phys_income_all) %>% 
  group_by(sex) %>% 
  summarise(mean_income = mean(income, na.rm = T))
```

### Mean income by height
Here you can see that on average, the taller you are, the more income you will get. This seems to be true for both female and male which was observed earlier in this analysis.
```{r}
height_income <- filter(phys_income_all) %>% 
 group_by(height) %>% 
 summarise(mean_income = mean(income, na.rm = T))
ggplot(height_income, aes(height,mean_income)) + geom_point() + geom_smooth(method = "lm")
```

### Mean income by height and gender
```{r}
filter(phys_income_all) %>% 
  group_by(height, sex) %>% 
  summarise(mean_income = mean(income, na.rm = T))
```

Here you can observe the difference of income between male and female at each height. You can generally see that as the heights get bigger, so does the income. It's also interesting to see the difference between incomes when a male and female are the same height.

Following our study of gender and height, we will begin to dive into our exploratory data analysis involving education and race.

## Study of Education

```{r}
library(tidyverse)
load("education_data_nlsy79.RData")
load("income_data_nlsy79.RData")
load("physical_data_nlsy79.RData")
educ_income <- education_data_nlsy79 %>% inner_join(income_data_nlsy79)
```

```{r}
ggplot(data = educ_income, aes(x=education)) + geom_bar()
```

Let's exclude the value 95, which means there are some ungraded education, from our analysis so that we can clearly see the histogram of education values between 0 and 25. 

```{r}
ggplot(data = filter(educ_income, education < 95), aes(x=education)) + geom_bar()

```


There are local peaks in the distribution at 12 and 16 years for high school and college degress, 14 years for associate degrees, and 18 years for an MBA or other 2-year master’s degree.\

Between 1981 and 2014 there are 58K cases of educational data. 

```{r}
sum(!is.na(educ_income$education))
```

### Explore the effect of Education on Income over time. 

We'll first look at the year 2014 and the distribution of income levels based on 4 education levels: High School, College, 2-year Masters, and 4-year post-undergraduate degree. There are some outliers above the $300K income level so in order to clearly see the differences, we removed the outliers. 

```{r}
ggplot(
  data = filter(educ_income, year==2014, education==12 | education==16 | education==18 | education==20, education < 95, income<300000), 
  aes(y=income)
) + geom_boxplot() + facet_wrap(~education)
```


It's clear that a college degree gives a significant boost in income over a high school degree with an increase in the median income of close to 25K. The upper quartile also saw a significant increase of nearly double in income. Masters and Post-graduate income levels are not that significantly different from college degree income levels.\


To understand how income levels change over time based on the education level the person achieved, we will be looking at how the mean income level per year for each education level behaves.

```{r}
mean_inc_educ_timeseries <- educ_income %>%
  filter(education == 12|education==16|education==18|education==20) %>%
  group_by(education, year) %>%
  summarise(mean_income = mean(income, na.rm=T))

ggplot(
  mean_inc_educ_timeseries, 
  aes(x=year,y=mean_income))+geom_point()+geom_smooth()+facet_wrap(~education)
```


This plot clearly demonstrates that with better education your income potential over time is much higher. We can see that with education 20 (4-year post-graduate degrees) in more recent times income has spiked, because presumably the degree is much more valuable and therefore can demand a higher income. Whereas with a high-school degrees, income levels stay relatively stagnant over time. 

### Mean Income by Education
```{r}
educ_income %>%
  filter(education < 95, education == 12|education==16|education==18|education==20) %>%
  group_by(education) %>%
  summarise(mean_income = mean(income, na.rm=T))
```

### Linear Model of Mean Income over time for High-School vs College
```{r}
summary(lm(year~mean_income, data=subset(mean_inc_educ_timeseries,education==12)))
summary(lm(year~mean_income, data=subset(mean_inc_educ_timeseries,education==16)))

```

## Study of Race

```{r}
library(tidyverse)
race_income <- physical_data_nlsy79 %>% inner_join(income_data_nlsy79) %>% subset(select=c("CASEID","year", "race","income"))
```
```{r}
ggplot(
  data = race_income, 
  aes(x=race)
) +geom_bar()
```

### Explore the effect of Race on Income over time. 

We'll first look at the year 2014 to initially determine how race influences income levels, removing the outliers of $300k income levels.  

```{r}
ggplot(
  data = filter(race_income, year==2014, income < 300000), 
  aes(y=income)
) + geom_boxplot() + facet_wrap(~race)
```


This plot indicates to us that in the year 2014, NBNH population earns higher incomes than Black/Hispanics. We can tell since the median and upper quartile of NBNH are noticeably higher than that of the other 2 races.\

Now we will study if being a particular race affect your income potential over time, by doing a time-series analysis of mean income levels by race. 

```{r}
mean_inc_race_timeseries <- race_income %>%
  group_by(race, year) %>%
  summarise(mean_income = mean(income, na.rm=T))

ggplot(
  mean_inc_race_timeseries, 
  aes(x=year,y=mean_income))+geom_point()+geom_smooth()+facet_wrap(~race)
```


This plot shows us that over time being Black or Hispanic, your income potential plateaus between 30k and 40k. Whereas for NBNH, income potential doesn't really seem to be limited at all. 

### Mean Income by Race

```{r}
race_income %>% group_by(race) %>% summarise(mean_income = mean(income, na.rm=T))
```
### Mean Income by Race and Year

```{r}
race_income %>% group_by(race, year) %>% summarise(mean_income = mean(income, na.rm=T))
```

## Study of Race and Education Combined
### Setting up the data
The next variable that we hope to explore is the relationship between education and race. As studied above, the variables of education and race are shown how they act independently but how do they effect one another. 
To begin, we must join the data sets of education and race to be able to study the impact of the two variables. 

```{r}
educ_race <- educ_income %>% inner_join(race_income)
```

This data frame will provide us with one main data frame to study the relationship between the two variables. 

Similar to the prior dataset, it is necessary to get rid of education values that are 95 or above as those values are clearly outliers.

```{r}
educ_race <- educ_race %>% filter(education <95)
```

Similarly, we will exclude all NA values for race as we cannot set a median value or any other estimate for a persons race in this data set.

```{r}
 sum(is.na(educ_race$race))
```

Using the function above, we can see that our data does not have any NA values for race which allows us to proceed. 

```{r}
sum(is.na(educ_race$education))
```

Similarly, we can see that for this new dataset, we have a corresponding race for each variable of education with no missing values present. 

### Visuals of the present data
```{r}
ggplot(data=educ_race, mapping=aes(x=year,y=education, color=race ))+geom_point()
```

```{r}
ggplot(data=educ_race, mapping=aes(x=education, fill=race))+geom_bar()
```

According to the data, it appears that NBNH participants of the NLSY79 survey make up the majority of the present values for all education levels. This can clearly be seen in the histogram above as the percentage of blue in comparison to the other two colors is much greater. 

Clearly there is some correlation between these two variable but what is the extent. Since race is a non numeric variable, let's use a linear model to understand the data better.


```{r}
model <-lm(education~race, data=educ_race)
summary(model)
```

According to the T-values there is a significant correlation between race and education. However, we want to better understand the relationship between income and race and eduation.

Let's begin by plotting out data.

```{r}
ggplot(data=educ_race, mapping=aes(x=education,y=income, color=race))+geom_point()
```

As shown in the data above, as education levels increase it appears that income increases as well. In terms of race, it appears that NBNH is currently found to have more years of education as opposed to those who are black or hispanic. Those who participated in the survey who are black do appear to have the second most frequent responses in terms of years of education greater than 12. These higher education levels directly correlate to greater income and are seen to be possessed more frequently by those who are NBNH and black as opposed to the hispanic survey participants. 

```{r}
model2<- lm(income~race + education, data=educ_race)
summary(model2)
library(car)
avPlots(model2)
```

These graphs simply show the predictor variable on the x-axis and the y-variable of income as the response variable. The blue lines shows the assocation between them. Here we can see that income is fairly consistent across races but has a slight postive correlation in terms of education.

### Mean Income of the Data
Over the years, the mean income for all races has steadily increased and can be seen to do so in the graphs below. 

```{r}
mean_inc_race_timeseries <- race_income %>%
  group_by(race, year) %>%
  summarise(mean_income = mean(income, na.rm=T))

ggplot(
  mean_inc_race_timeseries, 
  aes(x=year,y=mean_income,color=race))+geom_point()+geom_smooth()+facet_wrap(~race)
```

```{r}
educ_race2 <- educ_race %>% inner_join(mean_inc_educ_timeseries)
```

When studying the correlation of the mean income and education and race, we can see again that as education levels increase that mean income also increases. Similarly, the NBNH make up the largest percentage of the larger mean incomes seen below in each category of education from 12 years on. 


```{r}
ggplot(educ_race2, aes(x=mean_income, color=race))+geom_bar()+facet_wrap(~education)
```

For another viewpoint of the data, we can separate by race to study the density of mean income as related to education and compare the three graphs to make conclusions. 

```{r}
ggplot(educ_race2, aes(x=education,y=income, color=education))+geom_point()+facet_wrap(~race)
```


Similarly, when looking at the previous data, we can see an increase in mean income with an increase in education as well as an increase in the NBNH community as opposed to the Hispanic and black survey participants.

### Mean Income and the effects of Education and Race in 2014

Lastly, let's check our most recent data one more time in order to make some final conclusions.

```{r}
educ_race2014 <- educ_race %>% filter(year==2014)
educ_race2014.5 <- educ_race2014 %>% group_by(education,race) %>% summarise(mean_income = mean(income,na.rm=T))
educ_race2014.5
```

```{r}
ggplot(data=educ_race2014.5, mapping=aes(x=education,y=mean_income, color=race))+geom_point()+geom_smooth(se=F)
```

In the most recent years of the survey, we can see that there is clearly a increase in mean income as education levels increase. In 2014, the trend shows that Hispanic individuals of higher levels of education are trending to having a higher mean income than those who are NBNH and black. Those who are NBNH who make it past 8th grade do sharply increase their mean income as their level of education increases. In the black survery participants, it appears that for those who continue past an 8th grade education, they do increase their mean income but at a much less sharp rate as compared to their NBNH and hispanic survey counterparts. They also seem to plateau as the other two races increase as education continues.

# Conclusion
In conclusion, income can be seen to be effected by many factors according to the nlsy79 survey. The first variables studied were the height of the participants as well as their gender. According to the data gathered, there was a slight increase in mean income as height increased in both male and female survey participants. Male mean income was also significantly higher than their female counterparts.\ 

The next variables that were explored were education and race and the effects that the variables had on each other. As imagined, as education levels increased the mean income of the survey participants increased. This makes sense as with greater knowledge comes more responsibilities within the workplace and greater pay. In regards to race, it appeared that the NBNH consistently had the highest records of mean income as compared to the black and Hispanic survey participants. However in recent years, it appears that the mean income for the Hispanic survey participants is sharply growing. This could signal good things for the hispanic community in regard to securing high paying jobs going forward.

